### 2015-09-14 vmcai-2016-tex/approach.tex
Changes from Ben's non-pushed work

\section{Symbolic Initialization}
\label{sec:precise}
Although the case-splitting methods used in~\gsetxt{} can be applied
to guarded value sets, a direct implementation would run into the same
path explosion problems. This work presents a new initialization scheme which avoids the
nondeterminism introduced by~\gsetxt{}. Called \emph{\symtxt{}}, this scheme leverages the core idea in generalized symbolic execution
with lazy initialization, using on-the-fly reasoning to model a
black-box input heap during symbolic execution. Unlike~\gsetxt{}, \symtxt{} constructs a single
symbolic heap and polynomially-sized path condition for each control
flow path.

@@

The initialization rule in \figref{fig:symInit} is invoked whenever 
an uninitialized field is accessed. The interaction
with the solver in the definition of the rules is denoted by
$\mathbb{S}(\phi)$ which returns true if $\phi$ is satisfiable.  The
check for uninitialized references is determined by the function
$\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) =
\{\lp\phi\ \cfgnt{l}\rp\ ...\}$ which a returns guarded value set 
containing those locations where the field $\cfgnt{f}$ is uninitialized.
\[
\begin{array}{rcl}
\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) & = &\{ \lp\phi\ \cfgnt{l}\rp \mid \lp \phi\ \cfgnt{l}\rp  \in \cfgnt{L}\lp \cfgnt{r}\rp  \wedge \\ & &
 \ \ \exists \phi^\prime \lp \lp \phi^\prime\ \cfgnt{l}_\mathit{un}\rp  \in \cfgnt{L}\lp \cfgnt{R}\lp l,\cfgnt{f}\rp\rp \wedge \\ & &
 \ \ \mathbb{S}\lp \phi \wedge \phi^\prime \rp\rp\}\\
\end{array}
\]

In~\figref{fig:symInit}, given the uninitialized set $\Lambda$ for
field $f$, the $\mathrm{min}_l$ function returns
$(\phi_x\ \cfgnt{l}_x)$ based on a lexicographical ordering of
uninitialized locations in $\Lambda$ to make initialization
deterministic.  The set $\rho$ contains reference-location pairs that
represent potential aliases, where $\mathrm{isInit}\lp\rp$ ensures that the references are initialized. 
There are four cases encoded in the
symbolic heap: (i) $\theta_\mathit{null}$ represents the condition
where $\cfgnt{l}_\mathit{null}$ is possible, (ii)
$\theta_\mathit{new}$ represents the case where $\cfgnt{r}_f$ points
to a fresh location, (iii) each member of $\theta_\mathit{alias}$
restricts $\cfgnt{r}_f$ to a particular alias in $\rho$, and at the
same time, not alias any member of $\rho$ that was initialized earlier
than the current choice, and (iv) $\theta_\mathit{orig}$ implements
conditional initialization to preserve the original heap structure.
These sets are added into the heap on $\cfgnt{r}_f$ after the fields
for $\cfgnt{l}_f$ are initialized to $\cfgnt{r}_\mathit{un}$.

@@

There are two rewrite rules in~\figref{fig:fHeap}(a), one for reading
the value of a field (field-access) and the other for writing to a
field (field-write). Both rules first check that the operations can be
performed on a non-null location. The field-access rewrite rule
in~\figref{fig:fHeap} dereferences a field of type $\cfgnt{C}$ and
uses \figref{fig:symInit}, to get a new symbolic heap that is
initialized on the field $\cfgnt{f}$. The symbolic heap is further
modified to include a new stack reference pointing to the guarded value set
(possible values of the field) returned during the dereferencing; the
new stack reference is the return value from the operation.

>>>>

\section{Symbolic Initialization}
\label{sec:precise}
%Although the lazy initialization method used in~\gsetxt{} can be applied
%to guarded value sets, a direct case-splitting implementation would run into the same
%path explosion problems as~\gstetxt{}. A more desirable approach would be to create a new~\symtxt{} scheme that uses the state merging properties inherent to guarded value sets in order to avoid creating new paths during dereferencing operations. To accomplish this in a manner equivalent to~\gsetxt{}, uninitialized reference values must be initialized to guarded value sets that exactly reproduce the behavior of the individual~\gsetxt{} paths. Since~\gsetxt{} either chooses a new reference, a null reference, or an alias to an existing reference during lazy initialization, it seems obvious that these should be the locations represented in the guarded value set produced by~\symtxt{}. Although choosing the appropriate locations is straightforward, choosing the guards is more complicated. Since lazy initialization does not alter the path condition, it provides few clues as to what guards might be used. Thus, choosing the appropriate guards is the key obstacle in creating the desired~\symtxt{} scheme.
%
%If the exact nature of the guards is opaque, a few of the requirements may be divined by observing the properties of references. Since in a concrete heap, no reference points to two locations at the same time, the guards in any value set must be mutually exclusive. Also, since every reference points somewhere~\footnote{Even so-called dangling pointers can be thought of as pointing to unallocated memory.}, the set of guards must be collectively exhaustive. Furthermore, the fact that the input heap may contain a practically unbounded number of distinct references indicates that the value domain of the variables used in the guards should be similarly unbounded.

%We don't really care about the locations where the references are stored, only whether references alias or not.

This work presents a new initialization scheme which avoids the
nondeterminism introduced by~\gsetxt{}. Called \emph{\symtxt{}}, this scheme leverages the core idea in generalized symbolic execution
with lazy initialization, using on-the-fly reasoning to model a
black-box input heap during symbolic execution. Unlike~\gsetxt{}, \symtxt{} constructs a single
symbolic heap and polynomially-sized path condition for each control
flow path.

@@

In~\figref{fig:symInit}, given the uninitialized set $\Lambda$ for
field $f$, the $\mathrm{min}_l$ function returns
$(\phi_x\ \cfgnt{l}_x)$ based on a lexicographical ordering of
uninitialized locations in $\Lambda$ to make initialization
deterministic.  The set $\rho$ contains reference-location pairs that
represent potential aliases, where $\mathrm{isInit}\lp\rp$ ensures that the references are initialized. 
There are four cases encoded in the
symbolic heap. The first three correspond to the three types of choices made during lazy initialization: (i) $\theta_\mathit{null}$ represents the condition
where $\cfgnt{l}_\mathit{null}$ is possible, (ii)
$\theta_\mathit{new}$ represents the case where $\cfgnt{r}_f$ points
to a fresh location, (iii) each member of $\theta_\mathit{alias}$
restricts $\cfgnt{r}_f$ to a particular alias in $\rho$, and at the
same time, not alias any member of $\rho$ that was initialized earlier
than the current choice.

Unlike the first three cases, which correspond directly to~\gsetxt{} initialization rules, case (iv) is unique to~\symtxt{}. In this case, $\theta_\mathit{orig}$ implements conditional initialization to preserve the original heap structure. This step is necessary in order to maintain homomorphism\footnote{Homomorphism is an important property relied on by the correctness proof.} between symbolic heaps created using~\symtxt{} and the~\gsetxt{} heaps they are intended to represent. 
The sets from each of the four cases are added into the heap on $\cfgnt{r}_f$ after the fields
for $\cfgnt{l}_f$ are initialized to $\cfgnt{r}_\mathit{un}$.

@@

There are two rewrite rules in~\figref{fig:fHeap}, one for reading
the value of a field (field-access) and the other for writing to a
field (field-write). Both rules first check that the operations can be
performed on a non-null location. The field-access rewrite rule
in~\figref{fig:fHeap} dereferences a field of type $\cfgnt{C}$ and
uses \figref{fig:symInit}, to get a new symbolic heap that is
initialized on the field $\cfgnt{f}$. The symbolic heap is further
modified to include a new stack reference pointing to the guarded value set
(possible values of the field) returned during the dereferencing; the
new stack reference is the return value from the $\mathbb{VS}$ operation.

### 2015-09-14 vmcai-2016-tex/bisimulation.tex
Changes from Ben's non-pushed work

\section{Correctness}
\label{sec:bisim}
A major drawback of previous methods for creating test inputs for
symbolic heaps has been a lack of completeness. Incomplete heap
representations may not create a set of inputs that represents
the entire space of possible inputs. By missing some possible
inputs, the subsequent analysis may fail to recognize 
possible program behaviors.

The theorems in this section establish the soundness and
completeness of the symbolic heap approach with respect to Generalized
Symbolic Execution (\gsetxt{}). Intuitively, the theorems imply that any properties proven
with~\gsetxt{} can also be proved using the symbolic heap
algorithm. The proofs for the theorems, as well as
the complete set of semantic rules for~\gsetxt{} and~\symtxt{} can
be found in ~\cite{Hillery:2015}.

The proof shows the existence of a
bisimulation between sets of states related by~\gsetxt{} ($p \rgse
p^\prime$) and states related by the symbolic heap and update rules in this paper ($q \rsym
q^\prime$). The relations are on the universe of \emph{well-formed}
states $S$, which have the properties in \secref{sec:sh} with the
constraint that the states are \emph{feasible}: successors exist
unless at \cfgt{end}. Let $p \rgse p^\prime$ be a union over relations
for~\gsetxt{} (see~\cite{Hillery:2015}): $\rgse = \rgse^\mathit{A} \cup
\rgse^\mathit{A^\prime} \cup \rgse^\mathit{W} \cup
\rgse^\mathit{W^\prime} \cup \rgse^\mathit{E} \cup
\rgse^\mathit{E^\prime} \cup \rcom$, where \emph{A} is a field access
after evaluating the expression for the base reference $\lp
\cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp$, \emph{W} is
a field write after evaluating the expression for the right operand
$\lp \cfgnt{x}\ \cfgt{\$}\ \cfgnt{f}\ \cfgt{:=}\ \cfgnt{*} \rightarrow
\cfgnt{k}\rp$, \emph{E} is a reference compare after evaluating the
left and right operands $\lp \cfgnt{v}\ \cfgt{=}\ \cfgnt{*}
\rightarrow \cfgnt{k}\rp$, the prime symbol indicates a null reference
in the operation or a false outcome, and \emph{\com} is everything
else in the language. Any state relation, say $\rightarrow_x$, is
trivially extended to sets of states as
$$
P \hookrightarrow_x P^\prime \Longleftrightarrow \forall p \in P\ (\forall p^\prime\ (p \rightarrow_x p^\prime \Leftrightarrow p^\prime \in P^\prime))
$$
Let $\hookrightarrow_\gse$ be the extension of $\rgse$ to sets of
states. From this extension, a new meta transition relation is defined
over sets of states as
$$
\rsgse = \hookrightarrow_\gse^\mathit{A}
\cup \hookrightarrow_\gse^\mathit{A^\prime} \cup \hookrightarrow_\gse^\mathit{W} \cup
\hookrightarrow_\gse^\mathit{W^\prime} \cup \hookrightarrow_\gse^\mathit{E} \cup \hookrightarrow_\gse^\mathit{E^\prime}
\cup \hookrightarrow_\com
$$
The relation captures the notion of splitting groups of heaps at
certain operations. For example, suppose we have a set $P$ containing a single state $p$
with all references uninitialized. If $p$ is a field access state, it has two
potential successors in~\gsetxt{}: a non-null reference and a null reference. 
Thus, the $\rsgse$ relation has two successors and divides $P$ into the two outcomes.

>>>>

\section{Correctness}
\label{sec:bisim}
%A major drawback of previous methods for creating test inputs for
%symbolic heaps has been a lack of completeness. Incomplete heap
%representations may not create a set of inputs that represents
%the entire space of possible inputs. By missing some possible
%inputs, the subsequent analysis may fail to recognize 
%possible program behaviors.

The theorems in this section establish the soundness and
completeness of the symbolic heap approach with respect to Generalized
Symbolic Execution (\gsetxt{}). Intuitively, the theorems imply that any properties proven
with~\gsetxt{} can also be proved using the symbolic heap
algorithm. What follows is a brief description of the requisite terminology,
followed by the theorem statements. The proofs for the theorems, as well as
the complete set of semantic rules for~\gsetxt{} and~\symtxt{} can
be found in ~\cite{Hillery:2015}.

The theorems assert the existence of a
bisimulation between sets of states related by~\gsetxt{} ($p \rgse
p^\prime$) and states related by the symbolic heap and update rules in this paper ($q \rsym
q^\prime$). The relations are on the universe of \emph{well-formed}
states $S$, which have the properties in \secref{sec:sh} with the
constraint that the states are \emph{feasible}: successors exist
unless at \cfgt{end}. Let $p \rgse p^\prime$ be a union over relations
for~\gsetxt{} (see~\cite{Hillery:2015}): $\rgse = \rgse^\mathit{A} \cup
\rgse^\mathit{A^\prime} \cup \rgse^\mathit{W} \cup
\rgse^\mathit{W^\prime} \cup \rgse^\mathit{E} \cup
\rgse^\mathit{E^\prime} \cup \rcom$, where \emph{A} is a field access
after evaluating the expression for the base reference $\lp
\cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp$, \emph{W} is
a field write after evaluating the expression for the right operand
$\lp \cfgnt{x}\ \cfgt{\$}\ \cfgnt{f}\ \cfgt{:=}\ \cfgnt{*} \rightarrow
\cfgnt{k}\rp$, \emph{E} is a reference compare after evaluating the
left and right operands $\lp \cfgnt{v}\ \cfgt{=}\ \cfgnt{*}
\rightarrow \cfgnt{k}\rp$, the prime symbol indicates a null reference
in the operation or a false outcome, and \emph{\com} is everything
else in the language. Any state relation, say $\rightarrow_x$, is
trivially extended to sets of states as
$$
P \hookrightarrow_x P^\prime \Longleftrightarrow \forall p \in P\ (\forall p^\prime\ (p \rightarrow_x p^\prime \Leftrightarrow p^\prime \in P^\prime))
$$
Let $\hookrightarrow_\gse$ be the extension of $\rgse$ to sets of
states. From this extension, a new meta transition relation is defined
over sets of states as
$$
\rsgse = \hookrightarrow_\gse^\mathit{A}
\cup \hookrightarrow_\gse^\mathit{A^\prime} \cup \hookrightarrow_\gse^\mathit{W} \cup
\hookrightarrow_\gse^\mathit{W^\prime} \cup \hookrightarrow_\gse^\mathit{E} \cup \hookrightarrow_\gse^\mathit{E^\prime}
\cup \hookrightarrow_\com
$$
The relation captures the notion of splitting groups of heaps at
certain operations. For example, suppose a set $P$ contains a single state $p$
with all references uninitialized. If $p$ is a field access state, it has two
potential successors in~\gsetxt{}: a non-null reference and a null reference. 
Thus, the $\rsgse$ relation has two successors and divides $P$ into the two outcomes.
